A generalized definition of multidimensional item response theory parameters

### 2022-05-13 output/paper_draft/paper_M2PLM_parameters.Rmd
Expressions clarified after reviewing and updating mind map

A latent space where the traits are correlated is more readily interpretable as
a geometrical space with non-orthogonal axes,
often referred to as *general Cartesian coordinates* [@harman_modern_1970].
Such a space has the advantage that the cosines of the angles between two axes
can be interpreted as the correlation of the corresponding latent dimensions.
<!--# Reference needed? -->
<!-- Problem here: The latent space is not defined like that, according to
the results of section "Linear algebra of non-orthogonal coordinates". -->
The general case of MIRT model estimation implies estimating possibly
non-standard latent dimensions and non-null covariances among them,
a more realistic case than an identity matrix.
Assuming orthogonality among traits is usually unrealistic, be it in the
cognitive [e.g., @reckase_multidimensional_2009] or the non-cognitive domain<!--# Reference to example? -->.
Moreover, certain applications impose the estimation of
non-identity latent covariance matrices (e.g. multigroup equating) even when<!-- # Reference needed?. -->
interpreting the latent dimensions as substantive constructs is not a goal.
As we will argue, the usual definitions of the
MID and MDISC parameters do not generalize to the non-orthogonal geometry used
to represent correlated latent spaces.

>>>>

A latent space where the traits are correlated is easier to represent in
a geometrical space with non-orthogonal axes,
often referred to as *general Cartesian coordinates* [@harman_modern_1970].
Such a space has the advantage that the cosines of the angles between two axes
can be interpreted as the correlation of the corresponding latent dimensions.
<!--# Reference needed? -->
<!-- Problem here: The latent space is not defined like that, according to
the results of section "Linear algebra of non-orthogonal coordinates". -->
The general case of MIRT model estimation implies estimating possibly
non-standard latent dimensions and non-null covariances among them,
a more realistic case than an identity matrix.
Assuming orthogonality among traits is usually unrealistic, be it in the
cognitive [e.g., @reckase_multidimensional_2009] or the non-cognitive domain<!--# Reference to example? -->.
Moreover, certain applications impose the estimation of
non-identity latent covariance matrices (e.g. multigroup equating) even when<!-- # Reference needed?. -->
interpreting the latent dimensions as substantive constructs is not a goal.
As we will argue, the usual definitions of the
MID and MDISC parameters do not generalize to the non-orthogonal geometry used
to represent correlated latent spaces.
